Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-02 paper-2024-sac-marl-neighbouring-strategies.tex
draft q-leaerning (think where should we place it)

\usepackage{booktabs} % For formal tables
\usepackage{mytodonotes}

@@

where $\pi = \{\pi^i\}_{i \in N}$ represents the joint policy.  The goal in collaborative MARL is often to find a joint policy that maximizes this global average Q-value or a similar global objective.
%\subsection{Key Concepts}

@@

\paragraph{Centralized Actor-Critic Methods}

\subsection{Motivation}

\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
\begin{itemize}
\item Provide a detailed explanation of each neighbor-based method:
\begin{itemize}
\item \textbf{Independent learners:} Emphasize its simplicity and scalability but highlight its vulnerability to non-stationarity.
\item \textbf{NN-Averaging:} Describe the weight averaging process, discuss the influence of neighbor selection and averaging frequency, and mention weighted averaging variations.
\item \textbf{NN-Consensus:} Explain the selection of the best-performing network and its potential benefits and drawbacks.
\item \textbf{Experience Sharing:} Detail how agents share experiences, discuss asynchronous vs. synchronous sharing, and highlight its advantages in sparse interaction environments.
\end{itemize}
\end{itemize}
\section{Experimental Setup}\label{sec:experiments}
\begin{itemize}
\item Introduce the custom "collect the items" multi-agent environment:
\begin{itemize}
\item Describe the environment's dynamics, agent goals, and challenges.
\item Specify the action space (continuous and discrete) and observation space design.
\item Explain the reward structure, emphasizing the components that encourage cooperation.
\end{itemize}
\item Detail the implementation framework (Gymnasium, RLlib, PyTorch).
\item Specify the chosen RL algorithm (DQN) and its hyperparameter configuration.
\item Define the evaluation metrics: average episode length and average reward.
\end{itemize}

>>>>

\usepackage{booktabs} % For formal tables
\usepackage{graphicx}
\usepackage{autobreak}
\usepackage{mytodonotes}

@@

where $\pi = \{\pi^i\}_{i \in N}$ represents the joint policy. 
The goal in collaborative MARL is often to find a joint policy that maximizes this global average Q-value or a similar global objective.
There are several algorithms which try to find the optimal Q-value function, one of the most popular is the deep q-learning algorithm.
\subsubsection{Deep Q-Learning Formalization}
deep q-learning utilizes a deep neural network, 
often referred to as a Q-network, to approximate the action-value function $Q(s, a)$. 
This network takes the state $s$ as input and outputs a vector of Q-values, one for each possible action $a \in \mathcal{A}$.  
The optimal action-value function, denoted $Q^*(s, a)$, satisfies the Bellman optimality equation:

\[ Q^*(s, a) = \mathbb{E}_{s' \sim \mathcal{P}} \left[ r + \gamma \max_{a'} Q^*(s', a') | s, a \right] \]

where $\mathcal{P}$ is the state transition probability distribution.  
Deep Q-learning aims to train the Q-network to approximate $Q^*(s, a)$ by minimizing the loss function:

\[ \mathcal{L}(\theta) = \mathbb{E}_{(s, a, r, s') \sim \mathcal{D}} \left[ \left( r + \gamma \max_{a'} Q(s', a'; \theta^-) - Q(s, a; \theta) \right)^2 \right] \]

where $\theta$ are the parameters of the Q-network, $\theta^-$ are the parameters of a target Q-network (used for stability), 
and $\mathcal{D}$ is the experience replay buffer, a collection of past experiences $(s, a, r, s')$.
%
In the multi-agent setting, each agent $i$ has its own Q-network $Q^i(o^i, a^i; \theta^i)$ that takes its local observation $o^i$ as input.  
For centralized training, the global state $s$ can be used as input. 
The specific form of the multi-agent Q-function and the update rules depend on the chosen MARL algorithm (e.g., independent Q-learning, value decomposition methods).  
For example, in independent deep Q-learning, each agent updates its Q-network independently using its local reward $r^i$:
\begin{equation}
  \resizebox{\linewidth}{!}{$
  \mathcal{L}^i(\theta^i) = \mathbb{E}_{(o^i, a^i, r^i, o'^i) \sim \mathcal{D}^i} \left[ \left( r^i + \gamma \max_{a'^i} Q^i(o'^i, a'^i; \theta^{i-}) - Q^i(o^i, a^i; \theta^i) \right)^2 \right]
  $}
  \end{equation}
%\subsection{Key Concepts}

@@

\paragraph{Neighboring-Based Methods}
\paragraph{Centralized Actor-Critic Methods}

\subsection{Motivation}
MADRL faces a critical challenge in balancing scalability and performance.  
Centralized training methods, while effective, often suffer from computational bottlenecks and difficulties scaling to large numbers of agents. 
While distributed training approaches offer improved scalability, 
they frequently assume independent learners, 
neglecting crucial inter-agent coordination and leading to suboptimal global performance.  
%
Neighbor-based methods have shown promise in simpler, tabular settings by enabling localized coordination,
but their integration with deep reinforcement learning remains largely unexplored. 
This work bridges this gap by introducing a novel approach that combines the power of deep learning with the scalability of neighbor-based communication.  
Our method aims to achieve both efficient decentralized learning and effective coordination, enabling online adaptation and robust performance in complex multi-agent environments.
\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
\begin{itemize}
\item Provide a detailed explanation of each neighbor-based method:
\begin{itemize}
\item \textbf{Independent learners:} Emphasize its simplicity and scalability but highlight its vulnerability to non-stationarity.
\item \textbf{NN-Averaging:} Describe the weight averaging process, discuss the influence of neighbor selection and averaging frequency, and mention weighted averaging variations.
\item \textbf{NN-Consensus:} Explain the selection of the best-performing network and its potential benefits and drawbacks.
\item \textbf{Experience Sharing:} Detail how agents share experiences, discuss asynchronous vs. synchronous sharing, and highlight its advantages in sparse interaction environments.
\end{itemize}
\end{itemize}
\section{Experimental Setup}\label{sec:experiments}
\begin{itemize}
\item Introduce the custom "collect the items" multi-agent environment:
\begin{itemize}
\item Describe the environment's dynamics, agent goals, and challenges.
\item Specify the action space (continuous and discrete) and observation space design.
\item Explain the reward structure, emphasizing the components that encourage cooperation.
\end{itemize}
\item Detail the implementation framework (Gymnasium, RLlib, PyTorch).
\item Specify the chosen RL algorithm (DQN) and its hyperparameter configuration.
\item Define the evaluation metrics: average episode length and average reward.
\end{itemize}
